Data-based completion of grey-box models with neural networks

### 2020-09-23 systemID_NNs.tex
ok, working on the system id.

\section{Process modeling}
Consider a nonlinear plant evolving in continuous time as
\begin{align*}
  \dot x &= f_p(x, u) \\
  y &= C_px
\end{align*}
in which $x$ is the plant state (vector), $u$ is the manipulated control input,
and $y$ is the measurement. Assume that we don't have any 
disturbances/measurement and process noise 
for now, and we'll add those after we 
understand the issues in the nominal case.

\subsection{Completion of the u-y model}
In this class of problems, we assume that grey-box modeling is performed 
for the plant, and we wish to augment the grey-box model with neural networks
for improvement in the accuracy of the model. From first-principles modeling, 
we have the following (in continuous time, as grey-box modeling will 
almost always be in continuous time).
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= C_gx_g
\end{align*}

in which $x_g$ is some set of physical plant states chosen to be 
modeled. Denoted these physical states as $x_g=Cx$.
For observable nonlinear systems, 
we can construct the original plant state
using a history of measurements and control inputs.
Denote this history as 
\begin{align*}
  \mathbf{y}_{k:k-N} &= [y(k)', y(k-\Delta)', ..., y(k-N\Delta)']' \\
  \mathbf{u}_{k-1:k-N} &= [u(k-1)', u(k-2)', ..., u(k-N\Delta)']'
\end{align*}

There exists a $\phi(\cdot)$ and a number of delay 
time steps $N$, such that 
$x(k) = \phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})$.

We augment the grey-box model in continuous time 
as follows -
\begin{align*}
\dot x_g &= f_g(x_g, u) + 
f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N})\\
y &= C_gx_g 
\end{align*}
in which $f_{NN}(\cdot)$,
is a standard feedforward neural network. 
We assume that the measurements
are directly some subset of the physical grey-box states.

It is expected that the function $f_{NN}(\cdot)$ approximates
the difference between the true grey-box state dynamics
with what is modeled. 
\begin{align*}
  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \approx 
      f_p(x, u) - f_g(x_g, u) = 
      Cf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u)
\end{align*}

\subsection{Conversion to discrete time and training}

@@

>>>>

\section{Process modeling}
Consider a nonlinear plant evolving in continuous time as
\begin{align*}
  \dot x &= f_p(x, u) \\
  y &= h_p(x)
\end{align*}
in which $x$ is the plant state (vector), $u$ is the manipulated control input,
and $y$ is the measurement. Assume that we don't have any 
disturbances/measurement and process noise 
for now, and we'll add those after we 
understand the issues in the nominal case.

\subsection{Completion of the u-y model}
We assume that grey-box modeling is performed 
for the plant, and we wish to augment the grey-box model with neural networks
for improvement in the accuracy of the model. From first-principles modeling, 
we have the following grey-box model in continuous time
\begin{align*}
  \dot{x}_g &= f_g(x_g, u) \\
  y &= h_g(x_g)
\end{align*}

in which $x_g$ are the grey-box states, which are some subset 
of the plant states (denote $x_g = Gx$) chosen to 
be modeled. The dynamics of these grey-box states 
potentially have some missing terms 
in their ODEs.
For observable nonlinear systems, 
we can construct the original plant state
using a history of measurements and control inputs.
Denote this history as 
\begin{align*}
  \mathbf{y}_{k:k-N} &= [y(k)', y(k-\Delta)', ..., y(k-N\Delta)']' \\
  \mathbf{u}_{k-1:k-N} &= [u(k-1)', u(k-2)', ..., u(k-N\Delta)']'
\end{align*}

in which $\Delta$ is the sample 
time between two consecutive
prediction instants of the desired discrete time 
model to be used in MPC. There exists a $\phi(\cdot)$ 
and a number of delay 
time steps $N$, such that the plant state
at a given time can be written as
$x(k) = \phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})$.

We augment the grey-box model using 
neural networks as follows -
\begin{align*}
\dot x_g &= f_g(x_g, u) + 
f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N})\\
y &= h_g(x_g) + h_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})
\end{align*}
in which $f_{NN}(\cdot)$ and $h_{NN}(\cdot)$
are standard feedforward neural networks. We add the current 
control $u(k)$ as an input to the network $f_{NN}(\cdot)$ because 
there may be a mismatch in the actuator to the grey-box dynamics
as well.

It is expected that the functions $f_{NN}(\cdot)$ 
and $h_{NN}(\cdot)$ approximate
the following differences between the true grey-box state dynamics
and true measurement function with what is modeled.
\begin{align*}
  f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) &\approx 
      Gf_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}), u) - f_g(x_g, u) = 
      Gf_p(x, u) - f_g(x_g, u) \\
      h_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N}) &\approx 
      h_p(\phi(\mathbf{y}_{k:k-N}, \mathbf{u}_{k-1:k-N})) - h_g(x_g) = 
      h_p(x) - h_g(x_g) 
\end{align*}

\subsection{Conversion to discrete time and training}
We note that the network $f_{NN}(\cdot)$ is an addition 
to the grey-box dynamics in continuous time, whereas 
$h_{NN}(\cdot)$ is an addition directly to the 
measurement function. The Runge-Kutta-4 (RK4) method 
is used for the discretization as follows

\begin{align*}
  k_1 &= f_g(x_g(k), u(k)) + f_{NN}(\mathbf{y}_{k:k-N}, \mathbf{u}_{k:k-N}) \\
  k_2 &= f_g(x_g(k) + (\Delta/2)k_1, u(k)) + 
         f_{NN}(\mathbf{y}_{k+0.5:k-N+0.5}, 
                \mathbf{u}_{k+0.5:k-N+0.5})\\
  k_3 &= f_g(x_g(k) + (\Delta/2)k_2, u(k)) + 
                f_{NN}(\mathbf{y}_{k+0.5:k-N+0.5}, 
                      \mathbf{y}_{k+0.5:k-N+0.5})\\
  k_4 &= f_g(x_g(k) + \Delta k_3, u(k)) + 
                       f_{NN}(\mathbf{y}_{k+1:k-N+1}, 
                              \mathbf{u}_{k+1:k-N+1})\\
  x_g(k+\Delta) &= x_g(k) + (\Delta/6)(k_1 + 2k_2 + 2k_3 + k_4) \\
  \hat{y}(k+\Delta) &= h_g(x_g(k+\Delta)) + h_{NN}(\mathbf{y}_{k+\Delta:k-N+\Delta}, 
                                     \mathbf{u}_{k-1+\Delta:k-N+\Delta})
\end{align*}
in which the sequences $\mathbf{y}_{k+0.5:k-N+0.5}$ 
and $\mathbf{u}_{k+0.5:k-N+0.5}$ are defined as 
\begin{align*}
  \mathbf{y}_{k+0.5:k-N+0.5} &= 
    [y(k + \Delta/2)', y(k-\Delta/2)', ..., y(k-N\Delta + \Delta/2)']' \\
  \mathbf{u}_{k+0.5:k-N+0.5} &= 
  [u(k + \Delta/2)', u(k-\Delta/2)', ..., u(k-N\Delta + \Delta/2)']' 
\end{align*}
We assume a zero order hold on the control inputs between two consecutive 
MPC execution instants, such that $u(k + \Delta/2) = u(k)$. To get the 
measurements between two sample instants ($y(k + \Delta/2)$) for 
the RK4 predictions, a linear interpolation can be used 
or we can assume that additional 
measurements from the plant are available 
between two MPC execution instants as well.

The following optimization problem is solved to compute 
the weights in the neural networks 
\begin{align*}
 \underset{W_{fi}, b_{fi}, W_{hi}, b_{hi}}{\textnormal{min}} \sum_{i=0}^{N_s-1} 
 \dfrac{1}{2}(\hat{y}(i) - y(i))^2
\end{align*}
in which ($W_{fi}, b_{fi}, W_{hi}, b_{hi}$) are the weights in the 
two neural networks ($f_{NN}(\cdot)$ and $h_{NN}(\cdot)$), 
$\hat{y}$ are predictions of the augmented model, and $y$
are the measurements from the plant.
Note that this optimization requires an estimate 
of the initial grey-box state. We assume that 
before the open-loop system identification experiment 
performed for data collection, 
the plant is being operated at some 
steady-state such that input to output
responses of the plant and grey-box model are aligned with 
the use of an integrating disturbance model. We can use 
the grey-box state at that steady state as the initial condition 
for the above multi-step-ahead prediction-error
optimization.

@@

Consider the following plant, model,
and the expected function to be approximated 
with a neural network.

\textbf{Plant --}
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - k_2C_B + k_3C_C- \dfrac{C_B}{\tau}\\
  \dfrac{dC_C}{dt} &= k_2C_B - k_3C_C - \dfrac{C_C}{\tau}
\end{align*}
These ODEs represent the dynamics of the reaction system
$A \rightarrow B \rightleftharpoons C$ in a CSTR with constant 
height and temperature. The control input is $u = C_{A0}$, 
and the measurements are $y = [C_A, C_B]'$. The plant model 
is observable with these choices of the measurements and dynamics.

\textbf{Model --}
\begin{align*}
  \dfrac{dC_A}{dt} &= \dfrac{C_{A0} - C_A}{\tau} - k_1C_A\\
  \dfrac{dC_B}{dt} &= k_1C_A - \dfrac{C_B}{\tau}
\end{align*}
i.e, we assume that we don't know about the second reaction while 
performing grey-box modeling.
